---
title: "merpWS"
output: github_document
---
  
  <!-- README.md is generated from README.Rmd. Please edit that file -->
  
  ```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

##### Functions to access marine data via Web Services

An increasing number of marine data portals are making web services
available. These allow, alongside traditional downloads from the browser,
to access data programmaticely directly from scripting softwares such as Rstudio.

The main advantages are two-folds: 1) the data fetched by R scripts are always up to date,
a feature particularly useful when using data portals that are constantly added to; 2) 
fetching, formatting and cleaning steps can be done as part of a single seamless workflow.

merpWS is a collection of R functions making use of available marine web services to retrieve
data from the main marine data portals. 

## Installation
merpWS can be installed using the install_github function from the devtools package.  

```{r, echo = TRUE, eval = FALSE}
library(devtools)
devtools::install_github(repo = "MarineEcosystemResearchProgramme/merpWS")
library(merpWS)
```

## Usage
Below is the list of available functions:  
+ explore_cefas_data 

+ get_cefas_dataset_fields 

+ get_cefas_datasets  

Each function has a help file which can be accessed in R
by typing ? followed by the name of the function.  

## Development
merpWS functions are being developed openly on github. 