---
title: "merpWS"
output: github_document
---
  
  <!-- README.md is generated from README.Rmd. Please edit that file -->
  
  ```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

##### Functions to access marine data via Web Services

An increasing number of marine data portals are making web services
available. These allow, alongside traditional downloads from the browser,
to access data programmaticely directly from scripting softwares such as Rstudio.

The main advantages are two-folds: 1) the data fetched by R scripts are always up to date,
a feature particularly useful when using data portals that are constantly added to; 2) 
fetching, formatting and cleaning steps can be done as part of a single seamless workflow.

merpWS is a collection of R functions making use of available marine web services to retrieve
data from the main marine data portals. 

## Installation
merpWS can be installed using the install_github function from the devtools package.  

```{r, echo = TRUE, eval = FALSE, message = FALSE}
library(devtools)
devtools::install_github(repo = "MarineEcosystemResearchProgramme/merpWS")
library(merpWS)
```

```{r, echo = FALSE, eval = TRUE, message = FALSE}
library(merpWS)
```

## Usage
Below are possible workflows making use of merpWS functions.
These workflows exemplify how data from different data sources
can be combined to create "added data value". Typically we
combine an observed variable with one or several explanatory ones, 
thereby mimicking . Note that web services are in most cases
recent additions to data portals and bugs on their end are common. 
We are working in collaboration with data holding institutes to 
help them improve their offer. 

### Cefas data
CEFAS has recently made its API available. It allows exploring
and downloading some the data described in its new Data Hub.

```{r, echo = TRUE, eval = TRUE, cache = TRUE, message = FALSE, warning=FALSE}
# list all data files
mydatasets <- get_cefas_datasets()

# extract csv files (only downloadable format)
mydownloads <- mydatasets$to_download
mydata <- mydownloads[explore_cefas_data(mydownloads$Name, keyword = "SWT"), ]
```
```{r, echo = FALSE, eval = TRUE, cache = TRUE, message = FALSE, warning=FALSE}
# get some coastal temperature data
# 6 is problematic, 9 takes for ever, given up on it for the moment. all the others have worked!!
# save(res, file = "~/Desktop/all_cefas_temp.Rdata")
load('~/Google Drive/not_to_share/my_backup/myDocuments/postdoc/MERP_postdoc/temperature_data/all_cefas_temp.Rdata')
temp_data <- res[[1]]
for(i in 2:length(res)){
  if(!is.na(res[[i]])){
    temp_data <- rbind(temp_data, res[[i]])
  }
}
load('~/Google Drive/not_to_share/my_backup/myDocuments/github/emodnetMerpApp/data/uk_eez.Rdata')
library(merpWS)
library(robis)
library(ggmap)
library(mregions)
library(rgeos)
library(sp)
library(merpData)
library(magrittr)

load(file = '~/Google Drive/not_to_share/my_backup/myDocuments/postdoc/MERP_postdoc/temperature_data/gibbula_temp.Rdata')
load(file = '~/Google Drive/not_to_share/my_backup/myDocuments/postdoc/MERP_postdoc/temperature_data/stellatus_temp.Rdata')
load(file = '~/Google Drive/not_to_share/my_backup/myDocuments/postdoc/MERP_postdoc/temperature_data/montagui_temp.Rdata')
load(file = '~/Google Drive/not_to_share/my_backup/myDocuments/postdoc/MERP_postdoc/temperature_data/balanoides_temp.Rdata')
# same with Chthamalus stellatus (southern) and semibalanus balanoides (northern)
mydf <- data.frame(temp = c(stellatus_temp, balanoides_temp), 
                   species = c(rep("stellatus", length(stellatus_temp)), rep("balanoides", length(balanoides_temp))))
p <- ggplot(mydf) + geom_density(aes(temp, col = species)) + theme(legend.position = c(0.8, 0.9)) + xlab("Temperature")
```

CEFAS temperature data from different projects are available.
They can be downloaded one by one and aggregated. Below is an 
example for one of the data sets. 

```{r, echo = TRUE, eval = FALSE, message = FALSE}
temp_data <- download_cefas_data(recordset_id = mydata$Id[1])
```

We aggregate 10 CEFAS data sets together to obtain the
data.frame temp_data. The Dates are reformatted to create
year, month and day columns. 

```{r, echo = TRUE, eval = TRUE, message = FALSE}
# extract date in year, month and day from "Time"
library(magrittr)
inter <- strsplit(as.character(temp_data$Time), split = " ") %>%
  sapply(., function(x) x[[1]]) %>%
  strsplit(., split = "/") 
temp_data$day <- as.numeric(sapply(inter, function(x) x[[1]]))
temp_data$month <- as.numeric(sapply(inter, function(x) x[[2]]))
temp_data$year <- as.numeric(sapply(inter, function(x) x[[3]]))
```

```{r figure, fig.width=7, fig.height=6, fig.align = 'center', echo = TRUE, eval = TRUE, message = FALSE, warning=FALSE}
# plot temperature record locations. 
library(ggplot2)
library(ggmap)
mymap <- get_map(location = c(lon = -2.5, lat = 53.5), zoom = 5, maptype = "satellite")
ggmap(mymap) + geom_point(aes(x = Long, y = Lat), size = 0.5, data = temp_data)
```

```{r, echo = TRUE, eval = FALSE, cache = FALSE, message = FALSE}
# get some biotic data from obis for two species of barnacles with different thermal affinities: 
# the southern species Balanus stellatus and the northern species Semibalanus balanoides. 
library(robis)
library(ggmap)
library(mregions)
library(rgeos)
library(sp)
library(merpData)

uk_eez <- mr_shp("MarineRegions:eez", maxFeatures = NULL, filter = "United Kingdom Exclusive Economic Zone")
uk_eez <- SpatialPolygonsDataFrame(gSimplify(uk_eez, tol = 0.01, topologyPreserve = TRUE), 
data = uk_eez@data)

stellatus <- plot_obis("Chthamalus stellatus", shapefile = uk_eez, myresolution = 0.5, 
                     myzoom = 5, gridded = T)$obis_data
stellatus <- format_obis_date(stellatus)
stellatus <- stellatus[!is.na(stellatus$eventDate), ]

balanoides <- plot_obis("Semibalanus balanoides", shapefile = uk_eez, myresolution = 0.5, 
                     myzoom = 5, gridded = T)$obis_data
balanoides <- format_obis_date(balanoides)
balanoides <- balanoides[!is.na(balanoides$eventDate), ]

```

```{r, echo = TRUE, eval = FALSE, cache = FALSE, message = FALSE}
# draw realized thermal niche by matching occurrences with CEFAS temperature data. This can be a slow process
# that may take a few minutes depending on the size of both data sets to match.
stellatus_temp <- rep(NA, nrow(stellatus))
for(i in 1:length(stellatus_temp)) {
  stellatus_temp[i] <- match_env(env_lon = temp_data$Long, env_lat = temp_data$Lat, env_time = list(year = temp_data$year, month =     temp_data$month), bio_lon = stellatus$decimalLongitude[i], bio_lat = stellatus$decimalLatitude[i], bio_time = list(year = stellatus$year[i], month = stellatus$month[i]), max_dist = 100000, env_variable = temp_data$tC)
  print(stellatus_temp[i])
}

balanoides_temp <- rep(NA, nrow(balanoides))
for(i in 1:length(balanoides_temp)) {
  balanoides_temp[i] <- match_env(env_lon = temp_data$Long, env_lat = temp_data$Lat, env_time = list(year = temp_data$year, month =     temp_data$month), bio_lon = balanoides$decimalLongitude[i], bio_lat = balanoides$decimalLatitude[i], bio_time = list(year = balanoides$year[i], month = balanoides$month[i]), max_dist = 100000, env_variable = temp_data$tC)
  print(balanoides_temp[i])
}

mydf <- data.frame(temp = c(stellatus_temp, balanoides_temp), 
                   species = c(rep("stellatus", length(stellatus_temp)), rep("balanoides", length(balanoides_temp))))
ggplot(mydf) + geom_density(aes(temp, col = species))
```

```{r, echo = FALSE, eval = TRUE, fig.width=6, fig.height=5, fig.align = 'center', cache = FALSE, message = FALSE}
p
```

## Development
merpWS functions are being developed openly on github. 